FraCaS: Temporal Analysis

### 2020-06-23 paper2020/paper.tex
typography

{\small
\begin{verbatim}
s_279_1_p=
sentence 
(useCl past pPos
 (predVP
   (usePN (lexemePN "smith_PN"))
   (advVP
     (complSlash
      (slashV2a (lexemeV2 "write_V2")))
     (detCN (detQuant indefArt numSg)
       (useN (lexemeN "novel_N"))))
       (lexemeAdv "in_1991_Adv")))
s_279_3_h=
sentence
(useCl past pPos
 (predVP (usePN
   (lexemePN "smith_PN"))
   (advVP
     (complSlash
      (slashV2a (lexemeV2 "write_V2"))
      (usePron it_Pron))
     (lexemeAdv "in_1992_Adv"))))
\end{verbatim}

@@

Our dynamic and temporal semantics gives the following interpretation
for \verb!s_279_1_p! implies \verb!s_279_3_h!.
\begin{verbatim}
FORALL (fun a=>novel_N a)
(fun a=>(exists (b: Time),
((exists (c: Time),
(IS_INTERVAL Date_19910101 b /\
 IS_INTERVAL c Date_19911231 /\
 IS_INTERVAL b c /\
 appTime b c (write_V2 a)
   (PN2object smith_PN))))) ->
Not (exists (f: Time),
((exists (g: Time),
 (IS_INTERVAL Date_19920101 f /\
  IS_INTERVAL g Date_19921231 /\
  IS_INTERVAL f g /\
  appTime f g (write_V2 a) 
    (PN2object smith_PN)))))).
\end{verbatim}
In the above, one should remark the top-level quantification over the
novel (as explained in \cref{sec:our-semantics}), the quantification
over time intervals as individual timepoints, and the use of custom
operators for several constructions (\verb!FORALL!, \verb!Not!, \verb!IS_INTERVAL!,
\verb!appTime!). This use of custom operators is useful for several
generalisations (for example, we have quantifiers such as \verb!MOST! in
addition to \verb!FORALL!.)

Unfolding the definitions for these operators yield the following
proposition:
\begin{verbatim}
forall x : object,
novel_N x ->
(exists b c : Z,
   Date_19910101 <= b /\
   c <= Date_19911231 /\
   b <= c /\ write_V2 x SMITH b c) ->
(exists f g : Z,
   Date_19920101 <= f /\
   g <= Date_19921231 /\
   f <= g /\ write_V2 x SMITH f g) ->
False
\end{verbatim}
This is very close to our idealised representation of the problem
\cref{eq:smith-novel-1}. One difference is the use of abstract Coq
integers for timepoints. Using a discrete time allows us to use
predifined Coq tactics. The discrete nature of integers does not
interfere with the reasoning.

Finally, we can show a Coq proof for the above proposition:
\begin{verbatim}
Theorem  problem279 : Problem279aFalse.
cbv.
intros novel isSmithsNovel P1 H.
destruct P1 as
   [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
destruct H as
   [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
specialize writeUnique
  with (x := novel)(y := SMITH) as A.
unfold UniqueActivity in A.
specialize (A _ _ _ _ P1 H) as B.
lia.
Qed.
\end{verbatim}
}

@@

\section{Results and Evaluation}
\label{sec:results}
Our target is the FraCaS testsuite, which aims at covering a wide
range of common natural-language phenomena. The testsuite is structured in 9 sections:

The suite is structured according to the semantic phenomena involved
in the inference process for each example, and contains nine sections:
Quantifiers, Plurals, Anaphora, Ellipsis, Adjectives, Comparatives,
Temporal, Verbs and Attitudes.  The system described here focuses on
the Temporal section. However, it also supports the other eight
sections.  To our knowledge this is the first system which attempt to
target the temporal section. But in fact, our system is even provides
support for all the other section. A couple of decades after its
formulation, we propose a first attempt at covering the whole suite.
As such, there it is no other system to compare our system with, in
all aspects.  We can however compare with systems which target parts
of the FraCaS testsuite, as shown in \cref{tab:results}.

>>>>

\begin{small}
\begin{verbatim}
s_279_1_p=
sentence 
(useCl past pPos
 (predVP
   (usePN (lexemePN "smith_PN"))
   (advVP
     (complSlash
      (slashV2a (lexemeV2 "write_V2")))
     (detCN (detQuant indefArt numSg)
       (useN (lexemeN "novel_N"))))
       (lexemeAdv "in_1991_Adv")))
s_279_3_h=
sentence
(useCl past pPos
 (predVP (usePN
   (lexemePN "smith_PN"))
   (advVP
     (complSlash
      (slashV2a (lexemeV2 "write_V2"))
      (usePron it_Pron))
     (lexemeAdv "in_1992_Adv"))))
\end{verbatim}
\end{small}

@@

Our dynamic and temporal semantics gives the following interpretation
for \verb!s_279_1_p! implies \verb!s_279_3_h!.
{\small
\begin{verbatim}
FORALL (fun a=>novel_N a)
(fun a=>(exists (b: Time),
((exists (c: Time),
(IS_INTERVAL Date_19910101 b /\
 IS_INTERVAL c Date_19911231 /\
 IS_INTERVAL b c /\
 appTime b c (write_V2 a)
   (PN2object smith_PN))))) ->
Not (exists (f: Time),
((exists (g: Time),
 (IS_INTERVAL Date_19920101 f /\
  IS_INTERVAL g Date_19921231 /\
  IS_INTERVAL f g /\
  appTime f g (write_V2 a) 
    (PN2object smith_PN)))))).
\end{verbatim}
}
In the above, one should remark the top-level quantification over the
novel (as explained in \cref{sec:our-semantics}), the quantification
over time intervals as individual timepoints, and the use of custom
operators for several constructions (\verb!FORALL!, \verb!Not!, \verb!IS_INTERVAL!,
\verb!appTime!). This use of custom operators is useful for several
generalisations (for example, we have quantifiers such as \verb!MOST! in
addition to \verb!FORALL!.)

Unfolding the definitions for these operators yield the following
proposition:
{\small
\begin{verbatim}
forall x : object,
novel_N x ->
(exists b c : Z,
   Date_19910101 <= b /\
   c <= Date_19911231 /\
   b <= c /\ write_V2 x SMITH b c) ->
(exists f g : Z,
   Date_19920101 <= f /\
   g <= Date_19921231 /\
   f <= g /\ write_V2 x SMITH f g) ->
False
\end{verbatim}
}This is very close to our idealised representation of the problem
\cref{eq:smith-novel-1}. One difference is the use of abstract Coq
integers for timepoints. Using a discrete time allows us to use
predifined Coq tactics. The discrete nature of integers does not
interfere with the reasoning.

Finally, we can show a Coq proof for the above proposition:
{\small
\begin{verbatim}
Theorem  problem279 : Problem279aFalse.
cbv.
intros novel isSmithsNovel P1 H.
destruct P1 as
   [t0 [t1 [ct1 [ct2 [ct3 P1]]]]].
destruct H as
   [u0 [u1 [cu1 [cu2 [cu3 H]]]]].
specialize writeUnique
  with (x := novel)(y := SMITH) as A.
unfold UniqueActivity in A.
specialize (A _ _ _ _ P1 H) as B.
lia.
Qed.
\end{verbatim}
}

@@

\section{Results and Evaluation}
\label{sec:results}
Our target is the FraCaS testsuite, which aims at covering a wide
range of common natural-language phenomena. 
%
The suite is structured according to the semantic phenomena involved
in the inference process for each example, and contains nine sections:
Quantifiers, Plurals, Anaphora, Ellipsis, Adjectives, Comparatives,
Temporal, Verbs and Attitudes.  The system described here focuses on
the Temporal section. However, it also supports the other eight
sections.  To our knowledge this is the first system which attempt to
target the temporal section. But in fact, our system is even provides
support for all the other section. A couple of decades after its
formulation, we propose a first attempt at covering the whole suite.
As such, there it is no other system to compare our system with, in
all aspects.  We can however compare with systems which target parts
of the FraCaS testsuite, as shown in \cref{tab:results}.
